---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
set.seed(1014)
```

# purrrplus

The tidyverse makes working with data frames particularly easy. However, using pmap and running a function safely returns a list. purrrplus adds functionality to purrr that keeps analyis within the structure of a data frame, which is particularly useful for running simulations and specific types of data analysis tasks.

## Installation

```{r eval=FALSE, message=FALSE}
# install.packages("devtools")
devtools::install_github("stenhaug/purrrplus")
```

## Getting started

```{r message=FALSE}
library(purrrplus)
library(tidyverse) # most useful with the tidyverse
```

### Example

Imagine you have a function (which returns a named list or a named vector and might throw an error):

```{r}
calculate_if_positive <- function(a, b){
  if(a < 0 & b < 0) {stop("Both numbers are negative.")}
  else if(a < 0) {stop("Just the first number is negative")}
  else if(b < 0) {stop("Just the second number is negative")}
  
  list(add = a + b,
       subtract = a - b,
       multiply = a * b,
       divide = a / b)
}
```

And you want to apply this function to each row of a data frame (which might contain irrelevant variables):

```{r}
(numbers <- data_frame(a = c(-1, 0, 1, 2),
                      b = c(2, 1, 0, -1),
                      irrelevant = c("minneapolis", "st_paul", "minneapolis", "st_paul")))
```

### The old way

The irrelevant variable causes pmap to throw an error:

```{r error = TRUE}
output <- pmap(numbers, calculate_if_positive)
```

One way around this is to remove the irrelevant variables:

```{r}
numbers2 <- select(numbers, -irrelevant)
```

However, now calculate_if_positive function throws an error that stops everything if *any* of the rows contain a negative number:

```{r error = TRUE}
output <- pmap(numbers2, calculate_if_positive)
```

We are applying the function calculate_if_positive 4 times (once for each row in the data frame numbers). It should work for rows 2 and 3 and throw an error for rows 1 and 4. 

The purrrr function safely allows us to capture the results when it works and the error when it doesn't:

```{r}
output <- pmap(numbers2, safely(calculate_if_positive))
```

However, a function wrapped in safely returns a list which is difficult to work with:

```{r}
str(output)
```

### The purrrplus way

pmap_safely is the key function in purrrplus. pmap_safely takes a data frame (which might contain irrelevant variables) and applies a function (which returns a named list or a named vector and might throw an error) to each row. 

pmap_safely returns the data frame that was inputted but with an error and a result column which come from applying the function:

```{r}
(output <- pmap_safely(numbers, calculate_if_positive))
```

#### get_errors

get_errors allows for quick analysis of errors:

```{r}
get_errors(output)
```

get_errors with specific = TRUE breaks down the analysis by the specific error:

```{r}
get_errors(output, specific = TRUE)
```

get_results filters out rows with errors and unnests results such that each item in the list that the function returns has its own column:

#### get_results

```{r}
get_results(output)
```

Notice that "_result" is appended to each of these columns so any subsequent analysis can easily differentiate between variables the function produced and variables it didn't.
